---
title: "Tooth Growth"
output: 
    md_document

---
##Overview

This project analyzes the Tooth Growth data available in the R datasets package. It will provide a basic summary of the data, use confidence intervals and/or hypothesis testing to compare tooth growth, and will provide conclusions as well as the assumptions needed for those conclusions. The data set includes sixty observations showing the tooth length of 10 guinea pigs at each dosage of Vitamin C (0.5mg, 1mg, and 2mg) delivered either via orange juice or absorbic acid. Information on labels was found here: https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html. 

##Exploratory Analysis

First we load the data and take a look at it. 

```{r}
data(ToothGrowth)
str(ToothGrowth)
```

So we can see that there are 60 observations of three variables: len, supp, and dose. Len is the length of tooth, supp is the method of delivery (VC is absorbic acid and OJ is orange juice), and dose is the dosage of Vitamin C, either 0.5mg, 1mg, or 2mg. We can use the lattice library to get an overview of the data using the aggregate function. 

```{r}
library(lattice)
attach(ToothGrowth)
aggregate(len,list(dose,supp),FUN=function(x) c(x_mean=mean(x), x_sd=sd(x)))
detach(ToothGrowth)
```

We can also visualize some of this data in a box and whisper plot showing us length compared to dose for both absorbic acid and orange juice. 

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose) #remember to convert dose to a factor before proceeding
bwplot(len ~ dose | supp,data=ToothGrowth)
```

Based on this visualization, it looks like for both absorbic acid and orange juice, as the dose increases, the length of tooth increases. 

##Confidence Intervals and Hypothesis Testing

We want to investigate whether dosage (dose) or method of delivery (supp) are correlated to tooth length (len). We'll start by looking at the method of delivery. Our null hypothesis is that there is no difference in tooth length between the two methods. 

```{r}
t.test(len ~ supp, paired=F, var.equal=F, data=ToothGrowth)
```

We find a 95% confidence interval of -0.171, 7.571 for mean(OJ)-mean(VC). Our p value is 0.06063, which is not significant. Therefore, we would not reject the null hypothesis. 

So now we move on to dosage. For this, we need to set up some data frames to compare between the different dosages. 

```{r}
dose.one <- subset(ToothGrowth, dose %in% c(0.5, 1.0))
dose.two <- subset(ToothGrowth, dose %in% c(1.0, 2.0))
dose.three <- subset(ToothGrowth, dose %in% c(0.5, 2.0))
```

Then we'll run our t tests comparing the different data frames (note: it would be more appropriate to run an anova rather than multiple t tests, but for the purposes of this assignment, we'll use the t tests). Our null hypothesis is that there is no difference in tooth length between the different dosages. 

```{r}
t.test(len ~ dose, paired=F, var.equal=F, data=dose.one)
```

We find a 95% confidence interval of -11.98, -6.27 for mean(0.5)-mean(1.0) with a p value of 1.268e-07. Therefore, we reject the null hypothesis. 

```{r}
t.test(len ~ dose, paired=F, var.equal=F, data=dose.two)
```

We find a 95% confidence interval of -8.99, -3.73 for mean(1.0)-mean(2.0) with a p value of 1.906e-05. Therefore, we reject the null hypothesis. 

```{r}
t.test(len ~ dose, paired=F, var.equal=F, data=dose.three)
```

We find a 95% confidence interval of -18.15, -12.83 for mean(0.5)-mean(2.0) with a p value of 4.398e-14. Therefore, we reject the null hypothesis. 

##Conclusion and Assumptions

We have found that method of delivery does not result in a difference in tooth length, but that the dose does result in a difference in tooth length. This analysis assumes that these are independent populations, meaning that there must be a minimum of 60 guinea pigs being observed and no guinea pigs have been assigned to multiple groups. We are assuming that the guinea pigs represent an appropriate sample and that researchers taking these measurements were blinded to the dosages and method of delivery. We assume that there is no conflict of interest among the researchers. 